---
title: "Correlations between speech features and speaker characteristics / voice descriptions"
author: "Laura Fernández Gallardo"
date: "October 2017"
output: 
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = './')
knitr::opts_chunk$set(fig.width=10, fig.height=15, dpi=150)
```

```{r message=FALSE, warning=FALSE}

# clear
rm(list=ls())

# Libraries needed:

library(RCurl) # to read raw data from repo
library(caret) # pre-processing and feat imp
library(usdm) #  for variance inflation factor
library(mlr) # regression

```


## Objectives

Predicing voice descriptions (VD) from speech features, [extracted with OpenSMILE](https://github.com/laufergall/Subjective_Speaker_Characteristics/tree/master/data/speech_features).


## Load speech features

```{r}

path_github <- "https://raw.githubusercontent.com/laufergall/Subjective_Speaker_Characteristics/master/data/speech_features"

gemaps_m <- read.csv(text=getURL(paste0(path_github,"/eGeMAPSv01a_88_malespk.csv")), header=TRUE, sep=",")
gemaps_f <- read.csv(text=getURL(paste0(path_github,"/eGeMAPSv01a_88_femalespk.csv")), header=TRUE, sep=",")

```

## Ratings of voice descriptions

Subjective ratings were given to 20 speakers: 10 males and 10 females. 


```{r}

path_github2 <- "https://raw.githubusercontent.com/laufergall/Subjective_Speaker_Characteristics/master/data/subjective_ratings"

# load raw ratings
ratings <- read.csv(text=getURL(paste0(path_github2,"/VD_ratings.csv")), header=TRUE, sep=",")


# mean of ratings and split by gender
ratings_a <- aggregate(ratings[16:(16+33)], by=list(ratings$sample_heard), mean)
names(ratings_a)[1] <- 'sample_heard'

# merge ratings and features (10 samples only for each gender)
data_m <- merge(ratings_a, gemaps_m)
data_f <- merge(ratings_a, gemaps_f)

# just the 88 speech features
feats_m <- data_m[,(ncol(data_m)-88+1):ncol(data_m)]
feats_f <- data_f[,(ncol(data_f)-88+1):ncol(data_f)]

```


## Predicting: "melodisch" for male speakers

The item "melodisch" is relevant for the prediction of males' "mitfuehlend", "gelangweilt", "distanziert", "gleichgueltig", "freundlich" , "maennlich", and of females' "gleichgueltig" "ruhig" as seen [here](https://github.com/laufergall/Subjective_Speaker_Characteristics/blob/master/analysis/relationships_SC_VD/linearregression_VD_SC.md). 

Additionally, it seems to be easyly separable according to speakers' WAAT, as seen [here](https://github.com/laufergall/Subjective_Speaker_Characteristics/blob/master/analysis/voice_descriptions/explorative_analysis_ratings/analysis_ratings_cleanspeech.md)


### Pre-processing of speech features

Same as done [here](https://github.com/laufergall/Subjective_Speaker_Characteristics/tree/master/analysis/features_SC_VD).

Using the caret package: Box-Cox transformation to reduce skewness, centering and scaling.

```{r}

trans_m <- caret::preProcess(feats_m,
                            method = c('BoxCox','center','scale'))

transformed_m <- predict(trans_m, feats_m)

```


Calculate variance inflation factor (VIF) to deal with multicollinearity problems:

Naimi, B., Hamm, N. A. S., Groen, T. A., Skidmore, A. K. and Toxopeus, A. G.,"Where is Positional Uncertainty a Problem for Species Distribution Modelling?," Ecography, 37, pp. 191-203, 2014.

Identify collinear variables that should be excluded.

From the usdm package: "A VIF greater than 10 is a signal that the model has a collinearity problem".

```{r}

# vifcor
#v1 <- vifcor(transformed_m_2, th=0.5) 

# vifstep
v2_m <- vifstep(transformed_m, th=10)

transformed_m_v2 <- exclude(transformed_m,v2_m)

```


### mlr regression task

Predicting "melodisch" for male speakers.

```{r}

# just 10 observations and 6 features

mdata <- cbind(transformed_m_v2, data_m$melodisch)
names(mdata)[ncol(mdata)]<-'target'

# data for regression model

regrTask <- mlr::makeRegrTask(data = mdata, target = 'target')   

```


Function to perform regression with different learners and tuning hyperparameters (given as input parameters).

Learner fused with filter method for calculating the feature importance, based on linear correlation (as done [here](https://github.com/laufergall/Subjective_Speaker_Characteristics/blob/master/analysis/features_SC_VD/feature_importance_caret.md)). 

Leave-one-out cross-validation as resampling strategy, since we only have 10 observations.

I set the performance metric to be: median(abs(response - truth)).


```{r}

regr_fw_LOO_medae <- function(regr.lrn, ps) {

# learner fused with filter method for feature selection
lrn = mlr::makeFilterWrapper(learner = regr.lrn, fw.method = "linear.correlation")

# search space: going through different sizes of the feature subset 
# tuning by gridsearch
# resampling strategy: Leave-one-out Cv
# performance measure: medae

rdesc = mlr::makeResampleDesc("LOO")

res = mlr::tuneParams(lrn, 
                      task = regrTask, 
                      resampling = rdesc, 
                      par.set = ps, 
                      control = makeTuneControlGrid(),
                      measures = list(medae)
                      )

# train model with the optimal feature set found

lrn2 = makeFilterWrapper(learner = regr.lrn, 
                        fw.method = "linear.correlation", 
                        fw.abs = res$x$fw.abs)

mod = train(lrn2, regrTask)

print('selected features:')
cbind(getFilteredFeatures(mod))

}



# interesting plot to check correlations
# plotLearnerPrediction("regr.lm", features = "loudnessPeaksPerSec", task =regrTask)

```


Creating different learners and checking the regession performance.

#### regr.featureless

```{r}

regr.lrn = makeLearner("regr.featureless", method = 'mean')
# regr.lrn$par.set

# no features are selected for this regressor anyway
ps = makeParamSet(
  makeDiscreteParam("fw.abs", values = 1)
  )

regr_fw_LOO_medae(regr.lrn, ps)

```

#### regr.lm

Linear regression
 
Parameters: selecting a number of features 

```{r}

regr.lrn = makeLearner("regr.lm")
# regr.lrn$par.set

# parameters: selecting a number of features
ps = makeParamSet(
  makeDiscreteParam("fw.abs", values = seq(1, ncol(mdata)-1))
  )

regr_fw_LOO_medae(regr.lrn, ps)

```

 
#### regr.randomForest

Parameters: 
mtry: Number of variables randomly sampled as candidates at each split
ntree: Number of trees to grow

```{r}

regr.lrn = makeLearner("regr.randomForest")
regr.lrn$par.set

# parameters: selecting a number of features
ps = makeParamSet(
  makeDiscreteParam("fw.abs", values = seq(2, ncol(mdata)-1)),
  makeDiscreteParam("mtry", values = seq(1, 2)),
  makeDiscreteParam("ntree", values = seq(1, 10)) 
  )

regr_fw_LOO_medae(regr.lrn, ps)

```


#### regr.rknn

Random k-Nearest-Neighbors
Parameters: selecting a number of features and k

```{r}

regr.lrn = makeLearner("regr.rknn")
regr.lrn$par.set

ps = makeParamSet(
  makeDiscreteParam("fw.abs", values = seq(2, ncol(mdata)-1)),
  makeDiscreteParam("k", values = seq(1, 5))
  )

regr_fw_LOO_medae(regr.lrn, ps)

```

#### regr.rpart

Decision Tree

Parameters: selecting a number of feature, minsplit

```{r}

regr.lrn = makeLearner("regr.rpart")
regr.lrn$par.set

ps = makeParamSet(
  makeDiscreteParam("fw.abs", values = seq(1, ncol(mdata)-1)),
  makeDiscreteParam("minsplit", values = seq(1, 10))
  )

regr_fw_LOO_medae(regr.lrn, ps)

```


## Conclusions

* Performance (medae) on the regression task predicting subjective ratings given to "melodisch" for male speakers:

    * Baseline (selects mean): 16.8
    * Linear Regression: 11.0
    * Random Forest: 7.5 <- best performance, with just 2 selected features: "loudnessPeaksPerSec" and "equivalentSoundLevel_dBp"   
    * Random k-Nearest-Neighbors: 13.6
    * Decision Tree: 10.3
    
* This work contributes to the Lens model for the prediction of speaker characteristics given speech segments. The 2 steps are: 

    * from measurable acoustic speech parameters to voice desctiptions: this script shows that this can be done. Instead of "melodisch", we can also try to predict any of the other 33 subjective voice descriptions, such as "gepresst", "mit akzent", "unbetont", ... for male and for female speakers.
    * from voice descritions to speaker characteristics: there exist linear relationships that permit accurate predictions as shown [here](https://github.com/laufergall/Subjective_Speaker_Characteristics/blob/master/analysis/relationships_SC_VD/linearregression_VD_SC.md)
    
* Future work: more data needs to be collected to corroborate this, especially more subjective voice descriptions from other speakers. Here, we only have 20. The NSC database can be used to this aim.
