---
title: "Effects of Speaker Likability on Subjetive MOS"
author: "Laura Fernández Gallardo"
date: "December 2017"
output: 
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = './')
```


```{r echo=FALSE, message=FALSE, warning=FALSE}

# clear
rm(list=ls())

# paths
setwd("D:/Users/fernandez.laura/Documents/Work/Projects_Github/Subjective_Speaker_Characteristics/analysis/relationships_quality_WAAT")

path_github <- "https://raw.githubusercontent.com/laufergall/Subjective_Speaker_Characteristics/master/data/subjective_ratings"
  
 

#
path_data <- 'D:/Users/fernandez.laura/Documents/Work/WP8_SpkCharac_Quality/ListeningTest'
path_POLQA_data <- 'D:/Users/fernandez.laura/Documents/Work/WP8_SpkCharac_Quality/data_POLQA'
path_figures <- 'D:/Users/fernandez.laura/Documents/Work/WP8_SpkCharac_Quality/Figures'
path_R <- 'D:/Users/fernandez.laura/Documents/Work/WP8_SpkCharac_Quality/degrading_extremespeakers_8sentences/R'

# Libraries needed:

library(stringr) # for str_pad
library(ggplot2) # for plots
library(ez) # for ezANOVA
library(reshape2)  # for table, data manipulation
library(xtable) # for table for latex
library(knitr) # for kable
```


### Objectives

Subjective MOS test to study the effects of speakers' warmth-attractiveness (WAAT) on perceived transmitted speech quality.

### Speech Material

* 12 speakers: 6 males (3 with high WAAT, 3 with low WAAT) and 6 females (3 with high WAAT, 3 with low WAAT)
* Balanced content (8 sentences/speakers and no same wording so that the results are not dependent on one sentence)

* Degradations:
- downs8 + IRS + sv56 + c712 + G.711(at)64 + sv56
- downs8 + IRS + sv56 + c712 + AMR-NB(at)4.65 + sv56
- downs16 + HP135 + sv56 + P341 + G.722(at)64 + sv56
- downs16 + HP135 + sv56 + P341 + AMR-WB(at)6.6 + sv56
- downs32 + 14KBP + sv56 + SWV_EVS(at)16.4 + sv56
- MNRU 10 dB + sv56
- NB filtering 900 - 2500 Hz + sv56
- reference (48 kHz) + sv56

* Total: 8*12 = 96 stimuli (played randomly)
* mean length of the all selected 96 files: 8.7 s, sd = 1.1 s, min = 5.5 s, max = 12.0 s


### Listening test

```{r}

# speech degradations
degradations = c('G711','AMRNB','G722','AMRWB','EVS','filtered','MNRU','ref','BP900_2500','MNRU_10DB')  

# load raw answers
mydata <- read.csv2(text=getURL(paste0(path_github,"/MOS_Lik.csv")), header=TRUE, sep=",")  
mydata$MOS <- as.numeric(as.matrix(mydata$MOS))

# extract listener and stimulus information
mydata$listener <- substr(mydata$Participant,1,2)
mydata$listener_gender <- substr(mydata$Participant,4,4)
mydata$listener_age <- substr(mydata$Participant,6,7)

mydata$speaker <- substr(mydata$Filename,2,4)
mydata$speaker_gender <- substr(mydata$Filename,1,1)

mydata$distortion='ref'
for (i in 1:length(degradations)){
  pos_deg = grep(degradations[i] , mydata$Filename) 
  mydata$distortion[pos_deg]=degradations[i]
}

# add speakers' WAAT info (class)
spk_class <- read.csv(paste0(path_POLQA_data,'/spk_class.csv'))
spk_class$speaker <-  str_pad(spk_class$speaker, 3, pad = "0")
mydata <- merge(spk_class[,c(2,3)], mydata)

# get listeners' info
nstimuli = 96
table(mydata$listener_gender)/nstimuli
summary(as.numeric(mydata$listener_age))

# save these data
write.csv(mydata, paste0(path_data,'/subjective_MOS.csv'), row.names = F)

```

A group of 20 (10m, 10f) non-experts normal hearing listeners participated, aged 26.4 years on average (range: 19-34). Their mother tongue was German.


### Tables with MOS averaged

```{r}

# aggregate (mean across listeners and speakers) and reshape

mydata_agg <- aggregate(mydata$MOS, by = list(mydata$distortion, mydata$speaker_gender, mydata$class), mean)
names(mydata_agg) <- c('distortion','speaker_gender', 'WAAT', 'mean_MOS')

mean_MOS <- dcast(mydata_agg , distortion ~ WAAT + speaker_gender, value.var = "mean_MOS")

# sort from high to low MOS
mean_MOS$mean <- rowMeans(mean_MOS[,2:5])
mean_MOS <- mean_MOS[order(mean_MOS$mean, decreasing = T),]
rownames(mean_MOS) <- mean_MOS$distortion
mean_MOS <- mean_MOS[,-1]

# print table
options(digits=2)
# xtable(mean_MOS)
kable(mean_MOS[,c(1:4)])

write.csv(mean_MOS,paste0(path_data,'/mean_MOS_subjective.csv'))




## checking mean of high WAAT vs. mean of low WAAT

mean_high <- rowMeans(mean_MOS[,1:2])
mean_low <- rowMeans(mean_MOS[,3:4])

rbind(mean_high , mean_low)
mean_high - mean_low
min(mean_high - mean_low)
max(mean_high - mean_low)


## checking mean of males vs. mean of females

mean_m <- rowMeans(mean_MOS[,c(1,3)])
mean_f <- rowMeans(mean_MOS[,c(2,4)])

rbind(mean_m , mean_f)
mean_m - mean_f
max(mean_m - mean_f)





# Correlations of mean scores

mean_MOS_POLQA <- read.csv(paste0(path_R,'/mean_MOS_POLQA.RData'))


cor(mean_MOS_POLQA$high_m, mean_MOS$high_m)

cor(mean_MOS_POLQA$high_f, mean_MOS$high_w)

cor(mean_MOS_POLQA$low_m, mean_MOS$low_m)

cor(mean_MOS_POLQA$low_f, mean_MOS$low_w)

cor(mean_MOS_POLQA$mean, mean_MOS$mean)




## see correlations for all speakers of each class/gender, not aggregated MOS

mydata_agg2 <- aggregate(mydata$subjective_MOS, by = list(mydata$speaker, mydata$distortion, mydata$speaker_gender, mydata$class), mean)
names(mydata_agg2) <- c('speaker', 'distortion', 'gender', 'WAAT', 'subjective_MOS')

MOS_POLQA <- read.csv(paste0(path_R,'/MOS_POLQA.RData'))
MOS_POLQA <- MOS_POLQA[,c(2,9,3,4,7)]
MOS_POLQA$distortion <- as.character(MOS_POLQA$distortion)

# sort both datasets in the same way and merge
mydata_agg2 <- mydata_agg2[order(mydata_agg2$distortion),]
mydata_agg2 <- mydata_agg2[order(mydata_agg2$speaker),]

MOS_POLQA <- MOS_POLQA[order(MOS_POLQA$distortion),]
MOS_POLQA <- MOS_POLQA[order(MOS_POLQA$speaker),]

mydata_agg2$POLQA_MOS = MOS_POLQA$MOS

write.csv(mydata_agg2,paste0(path_data,'/Likability_MOS_POLQA_subjective.csv'), row.names = F)

# subsets:
subs <- mydata_agg2[mydata_agg2$gender=='w' & mydata_agg2$WAAT=='low',]
cor(subs$subjective_MOS, subs$POLQA_MOS)
sqrt(mean((subs$subjective_MOS - subs$POLQA_MOS)^2))
     

# all:
cor(mydata_agg2$subjective_MOS, mydata_agg2$POLQA_MOS)
sqrt(mean((mydata_agg2$subjective_MOS - mydata_agg2$POLQA_MOS)^2))



 





```

### Interaction plots

For each distortion separately, create interaction plot:

* view effects of:
- speaker gender (g)
- speakers' WAAT class (class)
- interaction (g:class)
* if lines are parallel: there is no interaction effect


```{r message = FALSE, warning=FALSE}

# rename "speaker_gender" to "g"
names(mydata)[names(mydata)=='speaker_gender'] = 'g'

# split for different distortions
mydata.split <- split(mydata, mydata$distortion)

# aggregate data
lapply(mydata.split, function(x){
  
  mydata_agg <- aggregate(x$MOS, by = list(x$g, x$class), mean)
  names(mydata_agg)[1]<-'g'
  names(mydata_agg)[2]<-'class' 
  names(mydata_agg)[3]<-'mean_MOS' 
  
  plot_jpg <- ggplot(x, aes(x = factor(class), y = MOS, fill = g)) + 
    geom_violin(adjust = .5, draw_quantiles = c(0.5)) + 
    geom_point(data = mydata_agg, aes_string(y = "mean_MOS", group = "g", color = "g")) + 
    geom_line(data = mydata_agg, aes_string(y = "mean_MOS", group = "g", color = "g")) + 
    theme_bw() +
    xlab("speaker WAAT class") +
    ylab("MOS") +  
    scale_y_continuous(limits = c(1, 5)) +
    theme(legend.position = "bottom")
  
  #jpeg(paste0(path_figures,"/violin_interaction_",x$distortion[1],".jpg"), family="Times") 
  #print(plot_jpg)
  #dev.off()

})



```

We observe that:

* AMR-NB: interaction effect: higher MOS for males compared to females, only for high WAAT. Low WAAT females get slightly better MOS compared to high WAAT. 
* G.711: lower MOS for low WAAT speakers. no gender effect. no interaction effect.

* AMR-WB: slightly lower MOS for low WAAT speakers. higher MOS for males. no interaction effect.
* G.722: interaction effect: higher MOS for males compared to females, only for high WAAT. Similar behaviour as for AMR-NB.

* BP900 - 2500 and MNRU 10 DB: lower MOS for low WAAT speakers. no or small gender effect. no or small interaction effect.

* EVS and ref: lower MOS for low WAAT speakers. higher MOS for males. no interaction effect.


### Two-way ANOVA

For each distortion separately, we perform 2-way repeated measures, to test the effects of: speaker gender (g), speakers' WAAT class (class), and interaction (g:class)

We use the package ez to compute ezANOVA: Bakeman, R. (2005). Recommended effect size statistics for repeated measures designs. Behavior Research Methods, 37 (3), 379-384. 


```{r warning= FALSE}

res <- sapply(mydata.split, function(x){
  
  output <- ezANOVA(data=x,
                    dv=MOS,
                    wid=listener,
                    within=.(g, class),
                    type = 2)
  
})

res

```




The effects observed in the interaction plots are confirmed with the ANOVA tests.

Significant effects (p<0.01) of speaker WAAT (class) or interactions (g:class) are found for: 

* AMR-NB (g, class, g:class) 
* BP (class)
* EVS (g, class)
* G.711 (class)
* G.722 (g, class, g:class) 
* MNRU (class) 
* ref (class)

Not found for: 

* AMR-WB (g)



Remarks:

* Effects of class observed: Consistently, lower MOS for low WAAT speakers and higher MOS for high WAAT speakers, for the same distortion. This manifests the importance of speakers' warmth-attractiveness for subjective speech quality assessments.
* Almost consistently, male speech is better rated than female speech for the same distortion
* The effect of speaker WAAT is even stronger than the effect of speker gender










