---
title: "Correlations between speech features and speaker characteristics / voice descriptions"
author: "Laura Fernández Gallardo"
date: "August 2017"
output: 
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = './')
knitr::opts_chunk$set(fig.width=5, fig.height=5, dpi=150)
```

```{r message=FALSE, warning=FALSE}

# clear
rm(list=ls())

# Libraries needed:

library(RCurl) # to read raw data from repo
library(caret) 
library(usdm)

```


## Objectives

Computing measures of feature importance for the prediction of:

* speaker characteristics (SC)
* voice descriptions (VD)


## Load speech features

```{r}

path_feats <- "https://raw.githubusercontent.com/laufergall/Subjective_Speaker_Characteristics/master/data/speech_features"

gemaps_m <- read.csv(text=getURL(paste0(path_github,"/eGeMAPSv01a_88_malespk.csv")), header=TRUE, sep=",")
gemaps_w <- read.csv(text=getURL(paste0(path_github,"/eGeMAPSv01a_88_femalespk.csv")), header=TRUE, sep=",")

gemaps <- rbind(gemaps_m, gemaps_w)
gemaps$id <- as.numeric(str_sub(gemaps$sample_heard,2,4))
gemaps$gender <- str_sub(gemaps$sample_heard,1,1)

gemaps <- gemaps[order(gemaps$id),]

```

male speakers: 126 observations.  89 variables = 88 features + filename

female speakers: 174 observations.  89 variables = 88 features + filename



## Feature importance for speaker characteristics

### Importance measures

### Plots

## Feature importance for voice descriptions

### Importance measures

### Plots



