---
title: "Factor analysis of voice descriptions"
author: "Laura Fernández Gallardo"
date: "September 2017"
output: 
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = './')

```


Perform factor analysis to reduce the 34-dimensional subjective voice descriptions to a smaller set of dimensions. Factor analysis instead of PCA in order to extract latent factors that represent the space of subjective voice descriptions that can be made by listening.


## 1) Load subjective ratings

Clear workspace.

```{r message=FALSE}

rm(list=ls())

```

Load necessary libraries.

```{r message=FALSE, warning=FALSE}

library(RCurl) # to read raw data from repo
library(GPArotation) 
library(psych) # for alpha analysis
library(knitr) # for kable

```

Set paths and read data.

```{r echo=FALSE}

setwd("D:/Users/fernandez.laura/Documents/Work/Projects_Github/Subjective_Speaker_Characteristics/analysis/voice_descriptions/factor_analysis")

path_github <- "https://raw.githubusercontent.com/laufergall/Subjective_Speaker_Characteristics/master/data/subjective_ratings"
  
data_raw <- read.csv(text=getURL(paste0(path_github,"/VD_ratings.csv")), header=TRUE, sep=",")  

```

## 2) Factor analysis for each speaker gender

Split data into male and female speakers to perform factor analysis separately, as males and females have different stereotypes. 

```{r}

data_split_g <- split(data_raw, data_raw$speaker_gender)
data_m <- data_split_g$m
data_f <- data_split_g$w

```

### 2a) Factor analysis of male speakers

```{r}

# always use z-scores
data_m <- data_m[order(data_m$listener_pseudonym),]
data_m<- cbind(data_m[,c(1, 5)], data.frame(do.call("rbind", as.list(by(data_m[,16:ncol(data_m)], data_m$listener_pseudonym, scale)))))
data_m <- data_m[order(data_m$listener_pseudonym),]

# find the no of factors
fa.parallel(data_m[,3:ncol(data_m)])

```

Parallel analysis suggests that the number of factors =  4  and the number of components =  3 . Hence, perform factor analysis with 4 factors:

```{r}
fa_m <- fa(data_m[,3:ncol(data_m)], nfactors=4 , rotate="oblimin", fm="minres") # this is default; no pricipal axis analysis
print(fa_m, cut=.2, digits=2, sort=T)

```

In order to only retain items with large main loading and small cross-loadings, we remove items when main loading <= .5  && (main loading - cross-loading) <= .2. We then run a second factor analysis with the retained items.

```{r}

## Second round: Run factor analysis again, removing items:

itemsremove <- c("heiser","kraftlos","unangenehm", "behaucht","natuerlich","ungleichmaessig","kurz","klanglos","laut","nicht_nasal","nicht_knarrend","langsam","mit_Akzent")
intemsremoveindexes <- match(itemsremove, colnames(data_m))

data_m_02 <- data_m[ ,-intemsremoveindexes]

names(data_m_02)

fa.parallel(data_m_02[,3:ncol(data_m_02)])
# Parallel analysis suggests that the number of factors =  4  and the number of components =  3 

fa_m_02 <- fa(data_m_02[,3:ncol(data_m_02)], nfactors=4 , rotate="oblimin", fm="minres") # this is default; no pricipal axis analysis
print(fa_m_02, cut=.2, digits=2, sort=T)

# fa.diagram(fa_m_02, cut=.40, digits=2)   # fa.graph for plotting (prior to alpha analysis)

```

This second factor analysis explained 54% of data variance.

Next, we examine Cronbach alphas to see if we should remove other items. 

```{r}

# Dimension 1 -  MR1
psych::alpha(as.data.frame(cbind(
  data_m_02$fest, data_m_02$praezise, data_m_02$unprofessionell, data_m_02$gleitend, data_m_02$zittrig, data_m_02$stockend, data_m_02$undeutlich)
),check.keys=TRUE)
# raw_alpha = 0.87 , no item removed


# Dimension 2 -  MR3
psych::alpha(as.data.frame(cbind(
  data_m_02$hart, data_m_02$gepresst, data_m_02$entspannt, data_m_02$warm, data_m_02$verbunden, data_m_02$schrill)
),check.keys=TRUE)
# raw_alpha = 0.79 , no item removed

# Dimension 3 -  MR4
psych::alpha(as.data.frame(cbind(
  data_m_02$monoton, data_m_02$auffaellig, data_m_02$melodisch, data_m_02$unbetont)
),check.keys=TRUE)
# raw_alpha = 0.82 , remove 2nd item
psych::alpha(as.data.frame(cbind(
  data_m_02$monoton, data_m_02$melodisch, data_m_02$unbetont)
),check.keys=TRUE)
# raw_alpha = 0.83


# Dimension 4 -  MR2
psych::alpha(as.data.frame(cbind(
  data_m_02$hell, data_m_02$hoch, data_m_02$scharf, data_m_02$glatt)
),check.keys=TRUE)
# raw_alpha = 0.80 , remove 4th item
psych::alpha(as.data.frame(cbind(
  data_m_02$hell, data_m_02$hoch, data_m_02$scharf)
),check.keys=TRUE)
# raw_alpha = 0.81

```


Compute new factor scores as a weighted average of the z-scores.

```{r}

loa_m <- fa_m_02$loadings
loa_m_df <- as.data.frame(loa_m[,])

# remove items from alpha analysis
remove_items <- c('auffaellig','glatt')
loa_m_df <- loa_m_df[-which(row.names(loa_m_df) %in% remove_items),]

# factor scores
fs_m_dim1 <- loa_m_df["fest",1]*data_m$fest + 
  loa_m_df["praezise",1]*data_m$praezise + 
  loa_m_df["unprofessionell",1]*data_m$unprofessionell + 
  loa_m_df["gleitend",1]*data_m$gleitend + 
  loa_m_df["zittrig",1]*data_m$zittrig + 
  loa_m_df["stockend",1]*data_m$stockend + 
  loa_m_df["undeutlich",1]*data_m$undeutlich

  
fs_m_dim2 <- loa_m_df["hart",2]*data_m$hart + 
  loa_m_df["gepresst",2]*data_m$gepresst + 
  loa_m_df["entspannt",2]*data_m$entspannt + 
  loa_m_df["warm",2]*data_m$warm +
  loa_m_df["verbunden",2]*data_m$verbunden + 
  loa_m_df["schrill",2]*data_m$schrill 

fs_m_dim3 <- loa_m_df["monoton",3]*data_m$monoton + 
  loa_m_df["melodisch",3]*data_m$melodisch + 
  loa_m_df["unbetont",3]*data_m$unbetont 

fs_m_dim4 <- loa_m_df["hell",4]*data_m$hell + 
  loa_m_df["hoch",4]*data_m$hoch +
  loa_m_df["scharf",4]*data_m$scharf

```

Write scores to file.

```{r}

factorscores_m <- data.frame("sample_heard"=data_m$sample_heard, "dim1"=fs_m_dim1, "dim2"=fs_m_dim2, "dim3"=fs_m_dim3, "dim4"=fs_m_dim4)

# average over raters
factorscores_m_averaged <- aggregate(factorscores_m[,2:ncol(factorscores_m)], by=list(factorscores_m$sample_heard), mean, na.rm=T)
names(factorscores_m_averaged)[1] <- "sample_heard"

write.csv(factorscores_m_averaged, "../../../data/generated_data/factorscores_VD_malespk.csv", row.names = F)

# save scores without averaging
factorscores_m_02 <- data.frame(data_m$listener_pseudonym, data_m$sample_heard,  "dim1"=fs_m_dim1, "dim2"=fs_m_dim2, "dim3"=fs_m_dim3, "dim4"=fs_m_dim4)

write.csv(factorscores_m_02, "../../../data/generated_data/factorscores_VD_malespk_notaveraged.csv", row.names = F)

```





### 2b) Factor analysis of female speakers

For female speakers, we follow a similar procedure as for male speakers. We also find 4 dimensions in the end, yet these are different latent factors as those found for male speech.


```{r}

# always use z-scores
data_f <- data_f[order(data_f$listener_pseudonym),]
data_f <- cbind(data_f[,c(1, 5)], data.frame(do.call("rbind", as.list(by(data_f[,16:ncol(data_f)], data_f$listener_pseudonym, scale)))))
data_f <- data_f[order(data_f$listener_pseudonym),]

# find the no of factors
names(data_f)[3:ncol(data_f)]
fa.parallel(data_f[,3:ncol(data_f)])
# Parallel analysis suggests that the number of factors =  5  and the number of components =  3 

fa_f <- fa(data_f[,3:ncol(data_f)], nfactors=5 , rotate="oblimin", fm="minres") # this is default; no pricipal axis analysis
print(fa_f, cut=.2, digits=2, sort=T)


```

Parallel analysis suggests that the number of factors =  5  and the number of components =  3.

In order to only retain items with large main loading and small cross-loadings, we remove items when main loading <= .5  && (main loading - cross-loading) <= .2.


```{r}

## Second round: Run factor analysis again, removing items:

itemsremove <- c("monoton","melodisch","praezise","klanglos","unbetont","kraftlos","mit_Akzent","kurz","entspannt","gepresst","verbunden","ungleichmaessig","natuerlich","scharf","nicht_knarrend","behaucht","langsam","nicht_nasal","laut")
intemsremoveindexes <- match( itemsremove, colnames(data_f))

data_f_02 <- data_f[ ,-intemsremoveindexes ]
names(data_f_02)[3:ncol(data_f_02)]

fa.parallel(data_f_02[,3:ncol(data_f_02)])
# Parallel analysis suggests that the number of factors =  4  and the number of components =  2 


fa_f_02 <- fa(data_f_02[,3:ncol(data_f_02)], nfactors=4 , rotate="oblimin", fm="minres") # this is default; no pricipal axis analysis
print(fa_f_02, cut=.2, digits=2, sort=T)

# fa.diagram(fa_f_02, cut=.40, digits=2)   # fa.graph for plotting (prior to alpha analysis)


```

This second factor analysis explained 53% of data variance.

Alpha analysis: Examine Cronbach alphas to see if we should remove other items. 


```{r}
 
# Dimension 1 -  MR1
psych::alpha(as.data.frame(cbind(
  data_f_02$zittrig, data_f_02$fest, data_f_02$gleitend, data_f_02$stockend)
),check.keys=TRUE)
# raw_alpha = 0.81, no item removed

# Dimension 2 -  MR2
psych::alpha(as.data.frame(cbind(
  data_f_02$hoch, data_f_02$hell, data_f_02$glatt, data_f_02$auffaellig, data_f_02$heiser)
),check.keys=TRUE)
# raw_alpha = 0.74 , remove 4th item
psych::alpha(as.data.frame(cbind(
  data_f_02$hoch, data_f_02$hell, data_f_02$glatt,data_f_02$heiser)
),check.keys=TRUE)
# raw_alpha = 0.76

# Dimension 3 -  MR4
psych::alpha(as.data.frame(cbind(
  data_f_02$unangenehm, data_f_02$undeutlich, data_f_02$unprofessionell)
),check.keys=TRUE)
# raw_alpha = 0.82, no item removed

# Dimension 4 -  MR3
psych::alpha(as.data.frame(cbind(
  data_f_02$schrill, data_f_02$hart, data_f_02$warm)
),check.keys=TRUE)
# raw_alpha = 0.71, no item removed



## Compute new factor scores as a weighted average of the z-scores

loa_f <- fa_f_02$loadings
loa_f_df <- as.data.frame(loa_f[,])

# remove items from alpha analysis
remove_items <- c('auffaellig')
loa_f_df <- loa_f_df[-which(row.names(loa_f_df) %in% remove_items),]

# factor scores
fs_f_dim1 <- loa_f_df["zittrig",1]*data_f$zittrig + 
  loa_f_df["fest",1]*data_f$fest + 
  loa_f_df["gleitend",1]*data_f$gleitend + 
  loa_f_df["stockend",1]*data_f$stockend 
  

fs_f_dim2 <- loa_f_df["hoch",2]*data_f$hoch + 
  loa_f_df["hell",2]*data_f$hell + 
  loa_f_df["glatt",2]*data_f$glatt + 
  loa_f_df["heiser",2]*data_f$heiser

  
fs_f_dim3 <- loa_f_df["unangenehm",3]*data_f$unangenehm + 
  loa_f_df["undeutlich",3]*data_f$undeutlich + 
  loa_f_df["unprofessionell",3]*data_f$unprofessionell
  
fs_f_dim4 <- loa_f_df["schrill",4]*data_f$schrill + 
  loa_f_df["hart",4]*data_f$hart + 
  loa_f_df["warm",4]*data_f$warm


```

Write scores to file.

```{r}


factorscores_f <- data.frame("sample_heard"=data_f$sample_heard, "dim1"=fs_f_dim1, "dim2"=fs_f_dim2, "dim3"=fs_f_dim3, "dim4"=fs_f_dim4)

factorscores_f_averaged <- aggregate(factorscores_f[,2:ncol(factorscores_f)], by=list(factorscores_f$sample_heard), mean, na.rm=T)
names(factorscores_f_averaged)[1] <- "sample_heard"

write.csv(factorscores_f_averaged, "../../../data/generated_data/factorscores_VD_femalespk.csv", row.names = F)


# save scores without averaging
factorscores_f_02 <- data.frame(data_m$listener_pseudonym, data_m$sample_heard,  "dim1"=fs_m_dim1, "dim2"=fs_m_dim2, "dim3"=fs_m_dim3, "dim4"=fs_m_dim4)

write.csv(factorscores_f_02, "../../../data/generated_data/factorscores_femalespk_VD_notaveraged.csv", row.names = F)


```


## 3) Summary of dimensions and loadings

* Male speech
Dimension 1: proficiency (precision / fluency) (*negative)
Dimension 2: tension
Dimension 3: melody
Dimension 4: brightness

* Female speech
Dimension 1: fluency (not connected to precision in contrast to male speech)  (*negative)
Dimension 2: brightness ( and also, smoothness with lower loading - can be neglected )
Dimension 3: proficiency precision (*negative)
Dimension 4: shrillness (related to tension dim of male speech)


### Male speakers

```{r}

# keep only English item translations
vd_items <- read.csv(text=getURL(paste0(path_github,"/VD_Questionnaire.csv")), header=TRUE, sep=",") 

loa_m_df$item <- rownames(loa_m_df)
loa_m_df_retained <- merge(loa_m_df, vd_items, by.x = 'item', by.y = 'right_German')
loa_m_df_retained <- loa_m_df_retained[,c(9,2:5)]

# set to NA the cross-loadings, to hide from kable
mean(loa_m_df_retained[,2:5][abs(loa_m_df_retained[,2:5])<0.4]) # avg of cros-loadings to be removed
loa_m_df_retained[,2:5][abs(loa_m_df_retained[,2:5])<0.4] <- NA

# dimension names 
names(loa_m_df_retained) <- c("item_English", "*neg proficiency precision",  "tension", "melody", "brightness")

# show loadings table with kable
options(knitr.kable.NA = '')
kable(loa_m_df_retained, digits = 2)


```



### Female speakers

```{r}

# keep only English item translations
loa_f_df$item <- rownames(loa_f_df)
loa_f_df_retained <- merge(loa_f_df, vd_items, by.x = 'item', by.y = 'right_German')
loa_f_df_retained <- loa_f_df_retained[,c(9,2:5)]

# set to NA the cross-loadings, to hide from kable
mean(loa_f_df_retained[,2:5][abs(loa_f_df_retained[,2:5])<0.4]) # avg of cros-loadings to be removed
loa_f_df_retained[,2:5][abs(loa_f_df_retained[,2:5])<0.4] <- NA

# dimension names
names(loa_f_df_retained) <- c("item_English","*neg fluency",  "brightness", "*neg proficiency precision", "shrillness")

# show loadings table with kable
options(knitr.kable.NA = '')
kable(loa_f_df_retained, digits = 2)

```

