---
title: "Applying BTL model to paired-comparison crowdsourcing data"
author: "Laura Fernández Gallardo"
date: "November 2016"
output: 
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = './')
knitr::opts_chunk$set(fig.width=5, fig.height=5, dpi=150)
```

Clear and set path.

```{r}

# clear
rm(list=ls())

path_github <- "https://raw.githubusercontent.com/laufergall/Subjective_Speaker_Characteristics/master/data/subjective_ratings/data_listeningtest8"
  

```


```{r message=FALSE, warning=FALSE}

# Libraries needed:

library(RCurl) # to read raw data from repo
library(eba) # BTL model
library(Hmisc) # for correlations

```

## Objectives

With data from a paired-comparison listening test conducted via crowdsourcing (Listening Test 8):

* Apply BTL model to obtain ratio-scaled preferences for voices
* Compare obtained values to those from Listening Test 7


## Preference matrix from crowdsourcing (Listening Test 8)

Preference matrix built from only accepted and automatically_accepted answers in crowdsroucing. See Matlab scripts. Row and column names are speaker pseudonyms.

```{r}

matrix_crowdee <- read.csv(text=getURL(paste0(path_github,"/preferencematrix_onlyaccepted.csv")), header=TRUE, sep=",")

rownames(matrix_crowdee) <- colnames(matrix_crowdee)

```


### Consistency checks

Following the same steps as in the paired-comparison [data analysis](https://github.com/laufergall/Subjective_Speaker_Characteristics/tree/master/analysis/voice_likability/paired_comparison/R) (Listening Test 7).

Concordance among listeners' preferences not applied to crowdsrouccing data.

Transitivity violations: 


```{r}

strans(matrix_crowdee)

```



### Scaling listeners' preferences

Since the BTL model has been shown to hold, utility scale (upsilon-scale) values were assigned to each speaker, representing their voice likability.

```{r}

btl1_crowdee <- eba(matrix_crowdee) # fit Bradley-Terry-Luce model
btl1_crowdee
# 	G2(91) = 88.31, p = 0.5605 it fits

```


Deriving utility scale values.

```{r}

# utility scale
usr_crowdee<-uscale(btl1_crowdee) #### no norm

# ordering from most liked to least liked
iir_crowdee<-order(usr_crowdee)

```


## Preference matrix from laboratory (Listening Test 7)

```{r}

path_github2 <- "https://raw.githubusercontent.com/laufergall/Subjective_Speaker_Characteristics/master/data/subjective_ratings/data_listeningtest7"
  
matrix_lab <- read.csv(text=getURL(paste0(path_github2,"/preferencematrix_total.csv")), header=TRUE, sep=",")

rownames(matrix_lab) <- colnames(matrix_lab)

# Fit BTL
btl1_lab <- eba(matrix_lab) # fit Bradley-Terry-Luce model
btl1_lab
# G2(91) = 66.08, p = 0.9772 it fits

# utility scale
usr_lab<-uscale(btl1_lab) #### no norm

# ordering from most liked to least liked
iir_lab<-order(usr_lab)


```

## Comparison of likability scores between Crowdee and lab

```{r}

cbind(usr_crowdee,usr_lab)
rp <- cor.test(usr_crowdee,usr_lab)
rp

# standard error
stderror_p <- sqrt((1-rp$estimate*rp$estimate)/(15-2)) 
stderror_p
# Pearson's correlation: r = 0.95, Standard error = 0.09




# Spearman's correlation:

rs <- rcorr(cbind(usr_crowdee,usr_lab), type = "spearman")
rs
stderror_s <- sqrt((1-rs$r[1,2]*rs$r[1,2])/(15-2)) 
stderror_s

```



